---
title: "Univariate Analysis"
date: "09/28/2022"
output: html_notebook
---


```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```


```{r}
GSS2016 <- read_csv("GSS2016.csv")
head(GSS2016)
```

***Central Tendancy and Variability***

```{r}
mean(GSS2016$age) 
GSS2016$age[GSS2016$age==99]=NA
mean(GSS2016$age, na.rm = TRUE) 
median(GSS2016$age, na.rm=TRUE)
```

```{r}
range(GSS2016$age, na.rm=TRUE)
max(GSS2016$age, na.rm = TRUE) - min(GSS2016$age, na.rm = TRUE)
quantile(GSS2016$age, na.rm=TRUE)
IQR(GSS2016$age, na.rm=TRUE)
```

```{r}
var(GSS2016$age, na.rm=TRUE)
sd(GSS2016$age, na.rm=TRUE)
## scale(GSS2016, center = TRUE, scale = TRUE)
## scale(GSS2016$age, center=FALSE, scale=TRUE)
```

```{r}
(72 - (mean(GSS2016$age, na.rm=TRUE)))/(sd(GSS2016$age, na.rm=TRUE))
```

```{r}
sample_n(GSS2016, 5, replace = FALSE)
```

```{r}
subset(GSS2016, sex==1)
```


***Univariate Analysis***

***Bar plot***

```{r}
values <- c(0.4, 0.75, 0.2, 0.6, 0.5)
barplot(values)
```

```{r}
group <- LETTERS[1:5]
barplot(values, names.arg = group)
```

***Stacked bar plot with legend***

```{r}
data <- as.matrix(data.frame(A = c(0.2, 0.4), B = c(0.3, 0.1), C = c(0.7, 0.1), D = c(0.1, 0.2), E = c(0.3, 0.3)))
rownames(data) <- c("Group 1", "Group 2")
barplot(data, col = c("#1b98e0", "#353436"))
legend("topright", legend = c("Group 1", "Group 2"), fill = c("#1b98e0", "#353436"))
```

***Grouped bar chart with legend***

```{r}
barplot(data, col = c("#1b98e0", "#353436"), beside = TRUE)
legend("topright", legend = c("Group 1", "Group 2"), fill = c("#1b98e0", "#353436"))
```

***Box plots***

```{r}
set.seed(8642)                                               
x <- rnorm(1000)
boxplot(x)
```

***Multiple box plots on one graph***

```{r}
y <- runif(1000)                                             
z <- rpois(1000, 3)
data <- data.frame(values = c(x, y, z), group = c(rep("x", 1000), rep("y", 1000), rep("z", 1000)))
head(data)
```

```{r}
boxplot(data$values ~ data$group, data, main = "My Boxplots", xlab = "My Boxplot Groups", ylab = "The Values of My Boxplots", col = c("red", "dark green", "purple"))
```



***Horizontal Box plot***

```{r}
boxplot(values ~ group, data, horizontal = TRUE, main = "My Boxplots", xlab = "My Boxplot Groups", ylab = "The Values of My Boxplots", col = c("red", "dark green", "purple"))
```

***Add Space Between Boxplots of Different Groups***

```{r}
data2 <- data                                                
data2$group <- c(rep("x1", 500), rep("x2", 500),rep("y1", 500), rep("y2", 500),rep("z1", 500), rep("z2", 500))
boxplot(values ~ group, data2, col = c("blue", "pink"),at = c(1, 2, 5, 6, 9, 10))
```

***Histograms***

***Non-Uniform Width***

```{r}
hist(rivers, main = "Length of Rivers in North America", col = "#1b98e0", breaks = c(0, 250, 300, 400, 500, 750, 1000, 2500, 5000))
```

***Manual number of breaks***

```{r}
hist(rivers, main = "Length of Rivers in North America", col = "#1b98e0", breaks = 50)
```

***Manual Axis limits***

```{r}
hist(rivers, col = "#1b98e0", xlim = c(0, 5000), ylim = c(0, 120))
```

***Overlaid Density Line***

```{r}
hist(rivers, col = "#1b98e0", ylim = c(0, 0.002), prob = TRUE)
lines(density(rivers), col = "red")
```


***Values on top of the bars***

```{r}
hist_values <- hist(rivers, ylim = c(0, 100))                                
text(hist_values$mids, hist_values$counts, labels = hist_values$counts, adj = c(0.5, - 0.5))
```


***Scatter plots***

***Fitting Line***

```{r}
set.seed(42424)                                   
x <- rnorm(500)
y <- x + rnorm(500)
plot(x, y, main = "This is my Scatterplot", xlab = "My X-Values", ylab = "My Y-Values")
abline(lm(y ~ x), col = "red")
```

***Smooth fitting line with the lowess function. Also modify color and point symbols***

```{r}
plot(x, y, main = "This is my Scatterplot", xlab = "My X-Values", ylab = "My Y-Values", pch = 16, col = "#1b98e0")                                       
lines(lowess(x, y), col = "green")
```


***Scatter plot with multiple groups***

```{r}
group <- rbinom(500, 1, 0.3) + 1
group_pch <- group                                
group_pch[group_pch == 1] <- 16
group_pch[group_pch == 2] <- 8
group_col <- group                                
group_col[group_col == 1] <- "red"
group_col[group_col == 2] <- "green"
plot(x, y, main = "This is my Scatterplot", xlab = "My X-Values", ylab = "My Y-Values", pch = group_pch, col = group_col)
legend("topleft", legend = c("Group 1", "Group 2"), col = c("red", "green"), pch = c(16, 8))
```

***Matrix of scatter plots***

```{r}
z <- runif(500)                                   
data <- data.frame(x, y, z)                       
pairs(data)
```


***Line plots***

```{r}
x <- 1:10                                            
y1 <- c(3, 1, 5, 2, 3, 8, 4, 7, 6, 9)
plot(x, y1, type = "l", main = "This is my Line Plot", xlab = "My X-Values", ylab = "My Y-Values", col = "blue", lwd = 4)
```

***Add points to line graph***

```{r}
plot(x, y1, type = "b", main = "This is my Line Plot", xlab = "My X-Values", ylab = "My Y-Values", col = "blue", lwd = 4)
```


***Plot multiple lines to one graph***

```{r}
y2 <- c(5, 1, 4, 6, 2, 3, 7, 8, 2, 8)                 
y3 <- c(3, 3, 3, 3, 4, 4, 5, 5, 7, 7)
plot(x, y1, main = "Line Plot", xlab = "My X-Values", ylab = "My Y-Values", type = "l")                                 
lines(x, y2, type = "l", col = "red")                 
lines(x, y3, type = "l", col = "green")
legend("topleft", legend = c("Line y1", "Line y2", "Line y3"), col = c("black", "red", "green"), lty = 1)
```


***Different point symbol for each line***

```{r}
plot(x, y1, main = "Line Plot", xlab = "X-Values", ylab = "Y-Values", type = "b", pch = 16)                      
lines(x, y2, type = "b", col = "red", pch = 15)
lines(x, y3, type = "b", col = "green", pch = 8)
legend("topleft", legend = c("Line y1", "Line y2", "Line y3"), col = c("black", "red", "green"), pch = c(16, 15, 8))
```


***Pie Graphs***

```{r}
df <- data.frame(group = c("Male", "Female", "Child"), value = c(25, 25, 50))
pie(df$value, labels = df$group, radius = 1, col = c("#999999", "#E69F00", "#56B4E9"))
```

***3-D Pie Charts***

```{r}
library(plotrix)
```


```{r}
plotrix::pie3D(df$value, labels = df$group, radius = 1.5, col = c("#999999", "#E69F00", "#56B4E9"))
```

***Explode the Pie Chart***

```{r}
plotrix::pie3D(df$value, labels = df$group, radius = 1.5, col = c("#999999", "#E69F00", "#56B4E9"), explode = 0.1)
```



