---
title: "Multivariate Regression"
date: "11/16/2022"
output: html_notebook
---


```{r}
library(Hmisc)
library(corrplot)
library(pastecs)
library(car)
```

```{r}
url <- "RStudio/GSS2016.csv"
GSS2016 <- read_csv(url)
head(GSS2016)
```


## Recoding the variables for education, maternal education, paternal education, hours of tv watched/day, and # of siblings, hours on the world wide web/week.


```{r}
GSS2016$educ[GSS2016$educ==98]=NA
GSS2016$educ[GSS2016$educ==99]=NA
GSS2016$maeduc[GSS2016$maeduc==97]=NA
GSS2016$maeduc[GSS2016$maeduc==98]=NA
GSS2016$maeduc[GSS2016$maeduc==99]=NA
GSS2016$paeduc[GSS2016$paeduc==97]=NA
GSS2016$paeduc[GSS2016$paeduc==98]=NA
GSS2016$paeduc[GSS2016$paeduc==99]=NA
GSS2016$tvhours[GSS2016$tvhours=='-1']=NA
GSS2016$tvhours[GSS2016$tvhours==98]=NA
GSS2016$sibs[GSS2016$sibs==98]=NA
GSS2016$sibs[GSS2016$sibs==99]=NA
GSS2016$wwwhr[GSS2016$wwwhr=='-1']=NA
GSS2016$wwwhr[GSS2016$wwwhr==998]=NA
GSS2016$wwwhr[GSS2016$wwwhr==999]=NA
```


## Compute new variable by adding together the years of education from the mother and the father

```{r}
GSS2016$mpeduc <- (GSS2016$maeduc + GSS2016$paeduc)
table(GSS2016$mpeduc)
```

## Generate correlation matrix

```{r}
cm2 <- GSS2016[, c("educ","mpeduc", "tvhours", "sibs", "wwwhr")] 
rcorr(as.matrix(cm2))
```

## Run Multiple Regression to see if any of the variables are associated with Education level

```{r}
reg2 <- lm(educ ~ mpeduc + tvhours + sibs + wwwhr, data = cm2)
reg2
```

```{r}
summary(reg2)
```

## Add some interaction objects to the linear model

```{r}
reg3 <- lm(formula = educ ~ mpeduc + tvhours + sibs + (mpeduc * tvhours) + (mpeduc * sibs) + (tvhours * sibs), data = cm2)
reg3
```

```{r}
summary(reg3)
```

## Produce a confidence interval for the coefficients of the Mulitple Linear Regression model

```{r}
confint(reg3)
```


## Call for the residuals

```{r}
resid(reg3)
```

## Make a histogram of the residuals to check for Normal Distribution

```{r}
hist(resid(reg3), col = "cornflowerblue")
```

Appears to be a Normal Distribution


## Check the Variance Inflation Factors (VIF) of the regression model without the interaction terms

```{r}
vif(reg2)
```
Values all close to 1, indication no issues of multicollinearity.




