---
title: "Mini-Project 1"
author: "Nash Lawrence"
date: "10/4/2022"
output:
  html_document:
    df_print: paged
---

This data set is about cars and their features.

Variable Description:

1.) reviews_count: Number of reviews on the car
2.) fuel_type: Type of fuel that the car uses (Petrol = 0, Diesel = 1, Other = 2)
3.) no_cylinder: number of cylinders contained in the car (0 for electrical cars)
4.) seating_capacity: number of people that can fit in the car
5.) transmission_type: type of transmission the car uses (Automatic = 0, Manual = 1, Electric = 2)
6.) body_type: body shape of the car (SUV = 0, Sedan = 1, Other = 2)
7.) max_power_bhp: maximum horsepower of the car


```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```



***1.) Load the data set and keep the 7 variables you want***

```{r}
url <- "CARS_2.csv"
carData <- read_csv(url)
carData <- carData %>% select(reviews_count, fuel_type, no_cylinder, seating_capacity, transmission_type, body_type, max_power_bhp)
carData
```


***2.) Convert the categorical variables to factors. Apply labels to the values of the variables***

```{r}
carData$fuel_type <- factor(carData$fuel_type, levels = c(0,1,2), labels = c("Petrol", "Diesel", "Other"))
carData$transmission_type <- factor(carData$transmission_type, levels = c(0,1,2), labels = c("Automatic", "Manual", "Electric"))
carData$body_type <- factor(carData$body_type, levels = c(0,1,2), labels = c("SUV", "Sedan", "Other"))
carData
```


***3.) Create two new "dummy" variables (0/1) from one continuous and one categorical variable***

```{r}
carData$autoCars <- ifelse(carData$transmission_type == "Automatic", 1, 0)
carData$over_fourSeats <- ifelse(carData$seating_capacity > 4, 1, 0)
carData$autoCars <- factor(carData$autoCars, levels = c(1,0), labels = c("Auto Trans", "Not Auto"))
carData$over_fourSeats <- factor(carData$over_fourSeats, levels = c(1,0), labels = c("More than 4", "4 or Less"))
carData
```

***4.) Table of Frequencies***

***a.) Find the cumulative frequency for all of your categorical variables (now totaling 5)***

```{r}
cbind(Freq = table(carData$fuel_type), Cum=cumsum(table(carData$fuel_type)))
cbind(Freq = table(carData$transmission_type), Cum=cumsum(table(carData$transmission_type)))
cbind(Freq = table(carData$body_type), Cum=cumsum(table(carData$body_type)))
cbind(Freq = table(carData$autoCars), Cum=cumsum(table(carData$autoCars)))
cbind(Freq = table(carData$over_fourSeats), Cum=cumsum(table(carData$over_fourSeats)))
```

***b.) Find the relative frequency (%) of each response category for your 5 variables***

```{r}
tblFuel <- table(carData$fuel_type)
propFuel <- round(prop.table(tblFuel), 3)*100
cbind(Freq = tblFuel, Perc = propFuel)

tblTrans <- table(carData$transmission_type)
propTrans <- round(prop.table(tblTrans), 3)*100
cbind(Freq = tblTrans, Perc = propTrans)

tblBody <- table(carData$body_type)
propBody <- round(prop.table(tblBody), 3)*100
cbind(Freq = tblBody, Perc = propBody)

tblDum1 <- table(carData$autoCars)
propDum1 <- round(prop.table(tblDum1), 3)*100
cbind(Freq = tblDum1, Perc = propDum1)

tblDum2 <- table(carData$over_fourSeats)
propDum2 <- round(prop.table(tblDum2), 3)*100
cbind(Freq = tblDum2, Perc = propDum2)
```


***5.) Central Tendency and Variation***

***a.) Find the mean, mode, and median of each of your 4 continuous variables***

```{r}
modeFunct <- function(x) {                      
  unique_x <- unique(x)
  tabulate_x <- tabulate(match(x, unique_x))
  unique_x[tabulate_x == max(tabulate_x)]
}
```

```{r}
modeFunct(carData$reviews_count)
median(carData$reviews_count)
mean(carData$reviews_count)
```

```{r}
modeFunct(carData$no_cylinder)
median(carData$no_cylinder)
mean(carData$no_cylinder)
```

```{r}
modeFunct(carData$seating_capacity)
median(carData$seating_capacity, na.rm = TRUE)
mean(carData$seating_capacity, na.rm = TRUE)
```

```{r}
modeFunct(carData$max_power_bhp)
median(carData$max_power_bhp)
mean(carData$max_power_bhp)
```


***b.)	Find the range, IQR, and SD for each of your 4 continuous variables***

```{r}
max(carData$reviews_count) - min(carData$reviews_count)
IQR(carData$reviews_count)
sd(carData$reviews_count)
```

```{r}
max(carData$no_cylinder) - min(carData$no_cylinder)
IQR(carData$no_cylinder)
sd(carData$no_cylinder)
```

```{r}
max(carData$seating_capacity, na.rm = TRUE) - min(carData$seating_capacity, na.rm = TRUE)
IQR(carData$seating_capacity, na.rm = TRUE)
sd(carData$seating_capacity, na.rm = TRUE)
```

```{r}
max(carData$max_power_bhp) - min(carData$max_power_bhp)
IQR(carData$max_power_bhp)
sd(carData$max_power_bhp)
```


***6.) Graphs***


***a.) Make a bar chart for a categorical variable. Include title, x/y labels, and a legend. Save it as PDF***

```{r}
fuelBar <- barplot(table(carData$fuel_type), main = "Number of Cars by Fuel Type", xlab = "Frequency", ylab = "Fuel Type", col = c("red", "dark green", "blue"))
legend("topright", legend = c("Petrol", "Diesel", "Other"), fill = c("red", "dark green", "blue"))
fuelBar
```

```{r}
pdf("rplot.pdf", width = 1350, height = 1350)
```



***b.)	Make a grouped box plot using a categorical and a continuous variable. Include title, x/y labels, legend. Save it as JPeg***

```{r}
transHorseBox <- boxplot(carData$max_power_bhp ~ carData$transmission_type, main = "Transmission Type vs Maximum Horsepower", xlab = "Type of Transmission", ylab = "Maximum Horsepower", col = c("red", "green", "purple"))
legend("top", legend = c("Automatic", "Manual", "Electric"), fill = c("red", "green", "purple"))
transHorseBox
```

```{r}
jpeg("rplot.jpg", width = 1350, height = 1350)
```


```{r}
myList <- list(data.frame(small = 1:10), data.frame(large = 11:20))
for(i in 1:2) {
  print (myList[1])
}
```



