---
title: "ST308 HW4"
author: "Nash Lawrence"
date: "11/21/2022"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(moments)
```


## 1.) Create a for loop to find the value of 10 factorial

```{r}
values <- 1
for (index in 1:10){
  values = values * index
}
paste("The factorial of 10 is:", values)
```


## Write a function that finds the factorial of a given number. Set default value to 10. Use an if statement to check if value of interest is greater than 100, if so stop the function and return a message saying the value is too large. If not, compute.


```{r}
facFunc <- function(val = 10){
  if (val >= 100){
    return("The input is too large.")
  }
  factor <- 1
  for (index in 1:val){
    if (val >= 100){
      break
    }
    else{
      factor = factor * index
    }
  }
  finalFac <- paste("The factorial of", val, "is:", factor)
  return(finalFac)
  
}
facFunc()
facFunc(8)
```


## 3.) Create a function 'convert_cm_to_in' that converts centimeters to inches. Determine how many inches a 25cm object is. 

```{r}
convertCM_toIN <- function(centimeter = 25){
  inches = centimeter * 0.39370079
  conValue <- paste("A", centimeter, "cm object is", inches, "inches.")
  return(conValue)
}
convertCM_toIN()
```

```{r}
facVec <- 0:100
convertCM_toIN(facVec)
```


## Add an addition input that rounds the result to two decimal places by default

```{r}
convertCM_toIN <- function(centimeter = 25){
  inches = centimeter * 0.39370079
  conValue <- paste("A", centimeter, "cm object is", round(inches, 2), "inches.")
  return(conValue)
}
facVec <- 0:100
convertCM_toIN(facVec)
```



## 5.) Create a function 'count_my_change' that adds up the total amount of money you input. 5 arguments: dollar, quarter, dime, nickel, penny. Default of all arguments set to zero. Calculate total by multiplying each value by its dollar value (Ex- 0.25 * quarters). Create a string saying "Your total is $amount" with the total, and put it inside a return statement.


```{r}
count_myChange <- function(dollar = 0, quarter = 0, dimes = 0, nickels = 0, pennies = 0){
  amount <- dollar + (0.25 * quarter) + (0.10 * dimes) + (0.05 * nickels) + (0.01 * pennies)
  total <- paste("Your total is $", amount)
  return(total)
}
count_myChange()
count_myChange(5, 3, 2, 1, 4)
```






